Syntactically Informed Natural Language Inference

### 2019-03-21 cse517/report/main.tex
Starting experiments section.

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2019}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{bm}
\usepackage{amsfonts}
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

The DA model is a simple and easily parallelizable approach to NLI.
We summarize it in this section so we can build upon it later.
The model
decomposes into attend, compare, and aggregate steps.

@@

\paragraph{Attend.} Compute unnormalized attention weights $e_{ij}$ with
a feed-forward neural network $F$:
\begin{align}
    e_{ij} := F(\bm{p})^\mathsf{T} F(\bm{h}).
\end{align}
Compute the aligned subphrases $\bm{H}i$ and $\bm{P}_j$:
\begin{align}
    \bm{H}_i &:= \sum_{j=1}^{\ell_h}\frac{\exp{(e_{ij})}}
                                        {\sum_{k=1}^{\ell_h}\exp{(e_{ik})}}h_j\,, \notag \\
    \bm{P}_j &:= \sum_{i=1}^{\ell_p}\frac{\exp{(e_{ij})}}
                                        {\sum_{k=1}^{\ell_h}\exp{(e_{ik})}}p_i\
\end{align}
where $\bm{H}_i$ is the weighted sum over $\bm{h}$ that aligns to $p_i$ and vice versa
for $\bm{P}_j$.

@@

We will now describe the different model architectures we experimented with.
In general, as the version number increases the complexity of the model will increase.

\paragraph{v1: Naive late fusion of syntactic features.}
The simplest way to incorporate $\bm{s}$ is to concatenate its final representation to
$\bm{v}^p$ and $\bm{v}^h$ as shown in figure~\ref{fig:v1}.
The prediction step in equation~\ref{eq:predict} then becomes:
\begin{align}
    \hat{\bm{y}} = H([\bm{v}^p, s^p_{\ell_p}, \bm{v}^h, s^h_{\ell_h}]).
\end{align}

@@

\paragraph{v2: Using syntactic features to attend.}
In the ``V3" model, we perform the same initial step of embedding the premise and 
hypothesis and expanding their dimensions (through a fully-connected layer) prior to 
attention. However, instead of calculating our similarity matrix using solely the 
embedded hypothesis and premise vectors as input, we also provide the encoded syntax 
parses of both inputs. The concatenation of these four vectors is used to derive the 
similarity matrix which is then utilized in the attention computations between the 
premise and hypothesis. \jbcomment{Insert V3.jpg here?}

With the ``V5" model we tried to incorporate the computed syntax of the inputs even 
earlier. Instead of concatenating the encoded syntax parses with the embedded inputs 
and using that as input for the computation of the similarity matrix, we concatenated 
the syntax parses with their respective inputs and passed that through the first 
fully-connected layer. The two resulting vectors representing the embedded inputs 
with syntactic information are then used to compute the similarity matrix. \jbcomment{Insert V5.jpg here?}

Our ``V7" model took an additional step of learning weights for the encoded syntax 
parses by passing the vectors through an additional fully-connected layer before 
introducing them into the rest of the network. This model was built on top of the 
``V3" model such that the weighted syntax parses are concatenated with the embedded 
premise and hypothesis statements when deriving the similarity matrix. \jbcomment{Insert V7.jpg here?}

\jbcomment{If we're only using data from the dependency parser we should probably 
remove the section following this comment}
Supplemental variations were done on top of these restructures of the network. One 
of the variations that we introduced was a retrained constituency parser. By 
retraining the constituency parser on the Penn Treebank dataset \citep{marcus1993building} 
we could change the dimensions of the encoded syntax vectors to better fit with our models. 
Additionally, we experimented with training our models using the dependency parser model 
derived by \citet{Dozat2016-gs} to see if the different types of encoded parses influenced 
the performance of the network.

@@

% \begin{figure}[h]
%   \includegraphics[width=0.45\textwidth]{v1}
%   \caption{A simple method of incorporating syntax into the decomposable
%     attention model.}
% \label{figure:v1}
% \end{figure}

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2019}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{bm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{multirow}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

The DA model is a simple and easily parallelizable approach to NLI.
We summarize it here so we can build upon it later.
The model
decomposes into attend, compare, and aggregate steps.

@@

\paragraph{Attend.} Compute unnormalized attention weights $e_{ij}$ with
a feed-forward neural network $F$:
\begin{align}\label{eq:attend}
    e_{ij} := F(p_i)^\mathsf{T} F(h_j).
\end{align}
Compute the aligned subphrases $\bm{H}i$ and $\bm{P}_j$:
\begin{align}
    \bm{H}_i &:= \sum_{j=1}^{\ell_h}\frac{\exp{(e_{ij})}}
                                        {\sum_{k=1}^{\ell_h}\exp{(e_{ik})}}h_j\,, \notag \\
    \bm{P}_j &:= \sum_{i=1}^{\ell_p}\frac{\exp{(e_{ij})}}
                                        {\sum_{k=1}^{\ell_h}\exp{(e_{ik})}}p_i\
\end{align}
where $\bm{H}_i$ is the weighted sum over $\bm{h}$ that aligns to $p_i$ and vice versa
for $\bm{P}_j$.

@@

We will now describe the different model architectures we experimented with.
In general, as the version number increases, the complexity of the model increases.

\paragraph{v1: Naive late fusion of syntactic features.}
The simplest way to incorporate $\bm{s}$ is to concatenate its final representation to
$\bm{v}^p$ and $\bm{v}^h$ as shown in figure~\ref{fig:v1}.

@@

The prediction step in equation~\ref{eq:predict} then becomes:
\begin{align}
    \hat{\bm{y}} = H([\bm{v}^p, s^p_{\ell_p}, \bm{v}^h, s^h_{\ell_h}])
\end{align}
where we concatenate the final hidden representation of the premise and hypothesis
when passed through the neural syntax parser with the aggregrated comparision vectors.
Everything else is identical to the DA model.

\paragraph{v2: Using syntactic features to attend.}
We wanted to experiment with incorporating syntactic features earlier in the model.
Instead of learning weights on the syntactic features, it is possible to
concatenate them to the projected input representations as show in figure~\ref{fig:v2}.

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/v2.png}
    \caption{Syntail v2 architecture.}
\label{fig:v2}
\end{figure}

Computing the
unnormalized attention weights in equation~\ref{eq:attend} then becomes:
\begin{align}
    e_{ij} := [F(p_i), s^p_i]^\mathsf{T} [F(h_j), s^h_i]
\end{align}
where instead of calculating the alignment using only the 
projected hypothesis and premise vectors, we now also use final hidden states of
the premise and hypothesis when passed through the neural syntax parser.
This seems rather naive since this model will not
learn any weights on the syntactic features, but in practice this model
performs very well. Everything else is identical to the DA model.

\paragraph{v3: Using syntactic features to project the premise and hypothesis.}
We decided to try incorporating the syntactic features even earlier in the model.
Instead of concatenating the final hidden states of the parsed premise and hypothesis
with the projected inputs, we concatenate the final hidden states of the parsed
premise and hypothesis with in the inputs immediately before projecting them. This
architecture is shown in figure~\ref{fig:v3}.

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/v3.png}
    \caption{Syntail v3 architecture.}
\label{fig:v3}
\end{figure}

Computing the unnormalized attention weights in equation~\ref{eq:attend} then becomes:
\begin{align}
    e_{ij} := F([p_i, s^p_i])^\mathsf{T} F([h_j, s^h_j]).
\end{align}
Everything else is identical to the DA model.

\paragraph{v4: Projecting the syntactic features before computing attention.}
After noticing a drop in performance from v2 to v3, we hypothesized that it was
due to not allowing our model to transform the syntactic features enough before
calculating attention.
We decided to learn weights to project the final hidden states of the parsed
premise and hypothesis by passing them through an additional fully-connected layer before 
concatenating them to the inputs. This architecture is shown in figure~\ref{fig:v4}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/v4.png}
    \caption{Syntail v4 architecture.}
\label{fig:v4}
\end{figure}

Computing the unnormalized attention weights in equation~\ref{eq:attend} then becomes:
\begin{align}
    e_{ij} := F([p_i, J(s^p_i)])^\mathsf{T} F([h_j, J(s^h_j)])
\end{align}
where $J$ is a feed-forward neural network.
Everything else is identical to the DA model.

% \jbcomment{If we're only using data from the dependency parser we should probably 
% remove the section following this comment}
% Supplemental variations were done on top of these restructures of the network. One 
% of the variations that we introduced was a retrained constituency parser. By 
% retraining the constituency parser on the Penn Treebank dataset \citep{marcus1993building} 
% we could change the dimensions of the encoded syntax vectors to better fit with our models. 
% Additionally, we experimented with training our models using the dependency parser model 
% derived by \citet{Dozat2016-gs} to see if the different types of encoded parses influenced 
% the performance of the network.

\subsection{Implementation Details}
\dpcomment{Add some hyper-parameter stuff here.}

@@

We compare our model's performance to the decomposable attention model
\citep{Parikh2016-em} and the ESIM model \citep{Chen2016-wl}.

\begin{table}[h]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{cccccc} 
\toprule
Model          & Embeddings      & Dev. Acc. & Test Acc.  \\ 
\midrule
Majority Class & n/a             & 63.3      & 60.3       \\
N-Gram         & n/a             & 65.0      & 70.6       \\
ESIM           & GloVe 840B 300d & 70.5      & 70.6       \\
DA             & GloVe 6B 300d   & 70.4      & 72.6       \\
DA             & ELMo            & 79.1      & 75.5       \\
\midrule
Syntail v2     & GloVe 6B 300d   & 79.3      & ?          \\
Syntail v3     & GloVe 6B 300d   & 82.3      & 77.4       \\
Syntail v5     & GloVe 6B 300d   & ?         & ?          \\
Syntail v7     & GloVe 6B 300d   & 81.3      & 77.0       \\
\bottomrule
\end{tabular}}
\label{table:experiments}
\caption{\dpcomment{think of a good caption, fix the version numbers}}
\end{table}

\subsection{Datasets}
For practical reasons, we only ran experiments on the SciTail dataset. SciTail (27k pairs) is
significantly smaller than SNLI (570k pairs) \citep{Bowman2015-is} or MultiNLI (433k pairs)
\citep{Williams2017-uh} which made it feasible to try many different architectures given
our limited resources.

Additionally, \citet{Gururangan2018-lj} showed that SNLI and MultiNLI
have significant annotation artifacts whereas SciTail is the first entailment set to be
completely derived from sentences that exist ``in the wild.''

\subsection{Baselines}

We list baseline performance reported by \citet{Khot2018-th} of a majority class predictor,
n-gram model, and ESIM \citep{Chen2016-wl} in table~\ref{table:experiments}. We reimplemented
and retrained the DA model and obtained similar test accuracies as reported by \citep{Chen2016-wl}.
Adding ELMo to the DA model improved test accuracy performance by approximately 3\%.
